---
title: "R Notebook"
output: html_notebook
---
2. The following R code is to inspect all 25 symbols:
 # A look at all 25 symbols 
df2 <- data.frame(x = 1:5 , y = 1:25, z = 1:25) 
s <- ggplot(df2, aes(x, y)) 
s + geom_point(aes(shape = z), size = 4) + sc?le_shape_identity() 
The output of the plot is given in the left panel. 
```{r}
df2 <- data.frame(x = 1:5 , y = 1:25, z = 1:25) 
s <- ggplot(df2, aes(x, y)) 
s+geom_point(aes(shape = z), size = 4) + scale_shape_identity() +geom_text(aes(label=z),hjust=-1,v?ust=-1)+ggtitle("Problem 1 : plot of 25 symbols\nby(Juber Rahman)")+theme_classic()+theme(axis.ticks=element_blank(),axis.text = element_blank(),axis.title = element_blank())


```

Use the code given below to get an output 
x <- as.data.frame.table(VADeat?s) 
colnames(x) <- c("age_group","sex_region","death") 
ggplot(x, aes(x = sex_region, y = death, fill = age_group)) + geom_bar(position = "dodge",stat = "identity") 
(a) Explain the effect of the first two commands. Why they are needed ? 
(b) Modify the pr?vious R code to i. add title with your name, ii. move the legend inside the plot, iii. write text within the bar with Mean= xx.xx, where "xx.xx" is the mean of each group in variable sex_region.
```{r}
p <-ggplot(x, aes(x = sex_region, y = death, fill = ag?_group)) + geom_bar(position = "dodge",stat = "identity")
p+ggtitle("Bar Diagram by Juber Rahman")

```

```{r}
p+ggtitle("Bar Diagram by Juber Rahman")+ theme(legend.position=c(1,1),legend.justification = c(1,.9))
```

```{r}
means <- aggregate(death ~ se?_region, x, mean)
ggplot(means, aes(x = sex_region, fill = sex_region, y = death)) +geom_bar(stat = "identity") +geom_text(aes(label = format(death, digits = 4), y = 0.7))


```

